### 2021-02-24 scripts/2178_chart_2.Rmd

---
title: "Untitled"
output: pdf_document
---
```{r}
citation("janitor")

```

```{r}
#### Workspace setup ####
library(here)
here()
#install.packages("tidyverse")
library(tidyverse) # used for data manipulation

#install.packages("ggpubr")
library(ggpubr) ##for combing figures in one graph

#install.packages("ggrepel")
library(ggrepel) #edit graphs
library(kableExtra)
#install.packages("janitor")
library(janitor) # Create tables
```
#import dataset
```{r}
simulated_dataset<-read_csv(here('outputs/survey/simulated_restaurant_data.csv'))

```

#Q3 layoff
```{r}
data.frame(table(simulated_dataset$q3_layoff,simulated_dataset$q1_city))%>%
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
    geom_bar(stat = "identity",position = position_dodge(), width = 0.7)+
   theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Number of Laid-offs")+
  ggtitle("Layoff Situation due to COVID 19")+
  guides(fill=guide_legend(title="City"))

```

#Q4 total sales
```{r}
#Peterborough
p_table <- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Peterborough",]$q4_total_sales))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

label_p<-paste(p_table$Var1,"\n",p_table$pct)%>%
  paste("%",sep = "")

pie(p_table$Freq,labels = label_p,col=c("#7FC97F","#BEAED4", "#FDC086", "#FFFF99", "#386CB0" ),
    main = "Peterborough: Total sales volume during Oct to Dec 2020 compared to July to Sept 2020",cex=0.6,cex.main=1)
```

```{r}
# use
q4_table <- data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))
data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1,"Prefer not to say", "Decreased more than 10%","Decreased 5% ~ 10%","Roughly unchanged","Increased 5% ~ 10%","Increased more than 10%")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Total Sale Change")+
  ggtitle("Total Sales Volume Change Oct to Dec vs July to Sept in 2020")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()
```

```{r}
#Brantford
b_table <- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Brantford",]$q4_total_sales))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

label_b<-paste(b_table$Var1,"\n",b_table$pct)%>%
  paste("%",sep = "")

pie(b_table$Freq,labels = label_b,col=c("#7FC97F","#BEAED4", "#FDC086", "#FFFF99", "#666666","#386CB0"),
    main = "Brantford: Total sales volume during Oct to Dec 2020 compared to July to Sept 2020",cex=0.5,cex.main=1)

```

# q5 total cost
```{r}
#Peterborough
p_table_cost <- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Peterborough",]$q5_total_cost))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

label_p<-paste(p_table_cost$Var1,"\n",p_table_cost$pct)%>%
  paste("%",sep = "")

pie(p_table_cost$Freq,labels = label_p,col=c( "#A6CEE3", "#1F78B4" ,"#B2DF8A" ,"#33A02C","#FB9A99"  ),
    main = "Peterborough: Total cost volume during Oct to Dec 2020 compared to July to Sept 2020",cex=0.6,cex.main=1)

```

```{r}
# use
q5_table <- data.frame(table(simulated_dataset$q5_total_cost,simulated_dataset$q1_city))
q5_table <-  q5_table %>% add_row(Var1 = "Decreased by more than 25%", Var2 = "Peterborough", Freq = 0) %>%
  add_row(Var1 = "Decreased by more than 25%", Var2 = "Brantford", Freq = 0)
q5_table%>%
  mutate(Var1 = fct_relevel(Var1,"Prefer not to say","Decreased by more than 25%", "Decreased by 5% ~ 25%","Roughly unchanged","Increased by  5% ~ 25%","Increased by more than 25%")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Total Cost Change")+
  ggtitle("Total Cost Volume Change Oct to Dec vs July to Sept in 2020")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()
```

```{r}
#Brantford
b_table_cost<- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Brantford",]$q5_total_cost))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

label_b<-paste(b_table_cost$Var1,"\n",b_table_cost$pct)%>%
  paste("%",sep = "")

pie(b_table_cost$Freq,labels = label_b,col=c("#A6CEE3", "#1F78B4" ,"#B2DF8A" ,"#FB9A99"  ),
    main = "Brantford: Total cost volume during Oct to Dec 2020 compared to July to Sept 2020",cex=0.6,cex.main=1)

```

# Q8 Survive

```{r}
# Use

data.frame(table(simulated_dataset$q8_survive,simulated_dataset$q1_city))%>%
    mutate(Var1 = fct_relevel(Var1 ,"Unknown","Longer than 6 months","3 to 6 months","Less than 3 months")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Expected Survival Time")+
  ggtitle("Expected Survival Time if Current Situation Continues")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()
```

```{r}
# No use

data.frame(table(simulated_dataset$q8_survive,simulated_dataset$q1_city))%>%
  ggplot(aes(Var1, Freq,fill=Var1)) +
    geom_bar(stat = "identity")+
  facet_wrap(~Var2)+
  scale_x_discrete(limits = c("Less than 3 months", "3 to 6 months", "Longer than 6 months","Unknown","Prefer not to say"))+
  ylab("Number of responses")+
  xlab("Expected survival time")+
  ggtitle("Expected Survival Time if Current Situation Continues")+
  theme_minimal()+
  theme(axis.text.x=element_text(size=rel(0.8), angle=20))

  #theme(axis.text.x = element_text(angle = 90))
```

# 10 recover
```{r}
# no use
q_10 <- data.frame(table(simulated_dataset$q10_recover,simulated_dataset$q1_city,simulated_dataset$q2_type))
data.frame(table(simulated_dataset$q10_recover,simulated_dataset$q1_city,simulated_dataset$q2_type))%>%
  ggplot(aes(Var1, Freq,fill=Var2)) +
    geom_bar(stat = "identity",position = position_dodge())+
  facet_wrap(~Var3)+
  ylab("Number of responses")+
  xlab("Number of months")+
  ggtitle("Expected Recovery Time Back to pre-COVID Situation")+
  guides(fill=guide_legend(title="City"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))+
   scale_x_discrete(limits = c("Six months or less","7 to 12 months","12 to 18 months","More than 18 months","Prefer not to say"))

```
```{r}
# use
data.frame(table(simulated_dataset$q10_recover,simulated_dataset$q1_city,simulated_dataset$q2_type))%>%
  mutate(Var1 = fct_relevel(Var1 ,"Prefer not to say","More than 18 months","12 to 18 months","7 to 12 months","Six months or less")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge())+
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Var3)+
  ylab("Number of responses")+
  xlab("Number of months")+
  ggtitle("Expected Recovery Time Back to pre-COVID Situation")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()

```

#q13 wage

```{r}
# brewer.pal(n = 8, name = "Dark2")
#Peterborough
p_table_wage <- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Peterborough",]$q13_CEWS))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

label_p<-paste(p_table_wage$Var1,"\n",p_table_wage$pct)%>%
  paste("%",sep = "")

pie(p_table_wage$Freq,labels = label_p,col=c( "#A6CEE3", "#1F78B4", "#B2DF8A" ),
    main = "Peterborough: Is your business receiving the Canada Emergency Wage Subsidy(CEWS)? ",cex=0.6,cex.main=1)

#Brantford
b_table_wage<- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Brantford",]$q13_CEWS))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

label_b<-paste(b_table_wage$Var1,"\n",b_table_wage$pct)%>%
  paste("%",sep = "")

pie(b_table_wage$Freq,labels = label_b,col=c("#A6CEE3", "#1F78B4","#FF7F00", "#B2DF8A" ),
    main = "Brantford: Is your business receiving the Canada Emergency Wage Subsidy(CEWS)? ",cex=0.6,cex.main=1)

```
```{r}
library(forcats)
```

# 14 loans
```{r}
# no use
data.frame(table(simulated_dataset$q14_loans,simulated_dataset$q1_city))%>%
  ggplot(aes(Var1, Freq,fill=Var2)) +
    geom_bar(stat = "identity",position = position_dodge())+
  ylab("Number of responses")+
  xlab("Grant value")+
  ggtitle("How much loans have you received for the COVID-19 related reasons?")+
  guides(fill=guide_legend(title="City"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 0))+
  coord_flip()

```
```{r}
?forcats::fact_reorder
```

```{r}
# use
data.frame(table(simulated_dataset$q14_loans,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1 ,"Prefer not to say","More than $50,000","$40,000 ~ $50,000","$20,000 ~ $40,000","$0 ~ $20,000")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
    geom_bar(stat = "identity", width = 0.8, position = position_dodge())+
  ylab("Number of responses")+
  xlab("Grant value")+
  ggtitle("How much loans have you received for the COVID-19 related reasons?")+
  guides(fill=guide_legend(title="City"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 0))+
  coord_flip()

```
```{r}
q_14 <- data.frame(table(simulated_dataset$q14_loans,simulated_dataset$q1_city))
```

# Q11

```{r}
simulated_dataset %>%
  tabyl(q1_city, q11_rate_support)%>%
  adorn_totals("row")%>%
  adorn_totals("col")%>%
  adorn_percentages("row")%>%
  adorn_pct_formatting(digits=2)%>%
  adorn_ns()%>%
  adorn_title("combined",row_name = "City", col_name = "Rating")%>%
  kable("html", align = 'ccccccc', caption = 'Number or percentage of each rating level')%>%
  kable_styling(full_width = F)
```